Creating the ``r params$package_name`` R package

### 2025-10-08 create-FSHybridPLS.Rmd
hybrid predictor object now has eval_point

The class definition is a simple list that contains a matrix and a collection of \code{fd} objects from the \code{fda} package, along with the necessary metadata.
- gram matrices are saved as a list.
```{r classdef:predictor_hybrid}
#' Create a predictor_hybrid object (S3 version, automatic basis size and Gram matrix)
#' Constructs a hybrid predictor object that stores both scalar and functional predictors,
#' and automatically computes Gram matrices (inner products of basis functions).
#'
#' @param Z A numeric matrix of dimension \code{n_sample × n_scalar} representing scalar predictors.
#' @param functional_list A list of functional predictors (e.g., \code{fd} objects from the \code{fda} package).
#'
#' @return An object of class \code{predictor_hybrid}, containing scalar and functional data with Gram matrices.
#' @export
predictor_hybrid <- function(Z, functional_list) {
  stopifnot(is.matrix(Z), is.numeric(Z))
  stopifnot(is.list(functional_list))

@@

  structure(
    list(
      Z = Z,
      functional_list = functional_list,
      gram_list = gram_list,
      gram_deriv2_list = gram_deriv2_list, # Store the new list
      n_basis_list = n_basis_list,
      n_sample = n_sample,
      n_functional = n_functional,
      n_scalar = n_scalar
    ),
    class = "predictor_hybrid"
  )
}
```

@@

```{r}
N <- 30
response <- c(1:N)
TRAIN_RATIO <- 0.7
create_mock_fd <- function(N_samples, seed) {
  set.seed(seed)
  mock_basis <- create.bspline.basis(rangeval = c(0, 1), nbasis = 5)
  C <- matrix(rnorm(5 * N_samples, mean = 10), nrow = 5)
  fd(C, mock_basis)
}
# 1. Setup Mock Hybrid Input
W_hybrid_mock <- predictor_hybrid(
  Z = matrix(1:(10*N), ncol = 10), # Simple sequential scalar data for verification
  functional_list = list(
    create_mock_fd(N, 1),
    create_mock_fd(N, 2)
  )
)

### DataSplitterNormalizer
R6 class to split functional and scalar predictors into train and test sets and apply normalization techniques including curve normalization, scalar normalization, and between-functional-scalar normalization.

```{r}
#'  Dataset Splitting and Normalization
#' @description 
DataSplitterNormalizer <- R6::R6Class(
  "DataSplitterNormalizer",

  public = list(

    data_original = NULL,  # Stores the original dataset
    n_sample = NULL,  # Number of samples
    train_ratio = NULL,  # Train-test split ratio
    idx_train = NULL,  # Train indices
    train=NULL,
    test=NULL,

    #' @title Initialize Data Splitter and Normalizer
    #' @description Constructs the instance, splits the data into train and test sets.
    #' Users must call `normalize_data()` separately.
    #' @param data_original List. Contains `functional_predictors` (list of `fd` objects)
    #' and `scalar_predictors` (matrix).
    #' @param train_ratio Numeric. Ratio of data to use for training (e.g., 0.8).
    #' @return An instance of `DataSplitterNormalizer`.
    #' @export
    initialize = function(data_original, train_ratio = 0.8) {
      self$n_sample <- length(data_original$response)
      self$train_ratio <- train_ratio
      self$data_original <- data_original

    },

    #' @title Split Data into Train and Test Sets
    #' @description Creates train/test indices and splits functional, scalar, and response variables.
    #' @return None. Updates train and test variables.
    #' @export
    split_data = function(idx_train = NULL) {
      # Use provided idx_train if available; otherwise, generate it
      if (is.null(idx_train)) {
        self$idx_train <- get_idx_train(self$n_sample, self$train_ratio)
      } else {
        self$idx_train <- idx_train
      }

      split_packed <- split.all(
        self$data_original$functional_1,
        self$data_original$functional_2,
        self$data_original$scalar,
        self$data_original$response,
        self$idx_train
      )

      # Assign train/test sets
      self$train <- list(
        functional_1 = split_packed$data_fun_1_train,
        functional_2 = split_packed$data_fun_2_train,
        scalar = split_packed$data_scalar_train,
        response = split_packed$response_train
      )
      self$test <- list(
        functional_1 = split_packed$data_fun_1_test,
        functional_2 = split_packed$data_fun_2_test,
        scalar = split_packed$data_scalar_test,
        response = split_packed$response_test
      )

    },

    #' @title Normalize Train and Test Data
    #' @description Applies curve, scalar, and between-functional-scalar normalization.
    #' @return None. Updates train and test sets.
    #' @export
    normalize_data = function() {
      # Initialize normalizers
      normalizer_functional <- normalizer_functional_predictor$new()
      normalizer_scalar <- normalizer_scalar_predictor$new()

      # Normalize first functional predictor
      normalizer_functional$normalize(self$train$functional_1, self$test$functional_1)
      self$train$functional_1 <- normalizer_functional$get_normalized()$train_normalized
      self$test$functional_1 <- normalizer_functional$get_normalized()$test_normalized

      # Normalize second functional predictor
      normalizer_functional$normalize(self$train$functional_2, self$test$functional_2)
      self$train$functional_2 <- normalizer_functional$get_normalized()$train_normalized
      self$test$functional_2 <- normalizer_functional$get_normalized()$test_normalized

      # Normalize scalar predictors
      normalizer_scalar$normalize(self$train$scalar, self$test$scalar)
      self$train$scalar <- normalizer_scalar$get_normalized()$train_normalized
      self$test$scalar<- normalizer_scalar$get_normalized()$test_normalized

      # Apply between-functional-scalar normalization
      self$normalize_between()

      # Normalize response variable
      self$normalize_response()
    },

    #' @title Normalize Between Functional and Scalar Predictors
    #' @description Scales scalar predictors to ensure comparability with functional predictors.
    #' @export
    normalize_between = function() {
      omega <- sum(diag(inprod(self$train$functional_1, self$train$functional_1))) +
        sum(diag(inprod(self$train$functional_2, self$train$functional_2)))

      omega <- omega / sum(self$train$scalar^2)

      self$train$scalar <- sqrt(omega) * self$train$scalar
      self$test$scalar <- sqrt(omega) * self$test$scalar
    },

    #' @title Normalize Response Variable
    #' @description Centers response variable based on training mean.
    #' @export
    normalize_response = function() {
      response_mean_train <- mean(self$train$response)
      self$train$response <- self$train$response - response_mean_train
      self$test$response <- self$test$response - response_mean_train
    } #must not end with ,

  )
)

@@

functions in MFPCA packages use an internal function calcBasisIntegrals. Somehow, this function did not work well on our simulations setting. Thus we modify the codes 
```{r}

###   Multivariate functional principal component analysis for functions on different (dimensional) domains

- This function calculates a multivariate functional principal component analysis (MFPCA) based on i.i.d. observations $x_1, \ldots, x_N$ of a multivariate functional data-generating process $X = (X^{(1)}, \ldots  X^{(p)})}{X = X^(1), \ldots, X^(p)$ with elements $X^{(j)} \in L^2(\mathcal{T}_j)}{X^(j) in L^2(calT_j)$ defined on a domain $\mathcal{T}_j \subset IR^{d_j}}{calT_j of IR^{d_j}$.
- In particular, the elements can be defined on different (dimensional) domains.
- The results contain the mean function, the estimated multivariate functional principal components \eqn{\hat \psi_1, \ldots, \hat \psi_M} (having the same structure as $x_i$), the associated eigenvalues $\hat \nu_1 \geq \ldots \geq \hat \nu_M > 0$ and the individual scores \eqn{\hat \rho_{im} = \widehat{<x_i, \psi_m>}}{\hat \rho_{im} = \hat{<x_i, \psi_m>}}.

- Univariate Expansions
    - The multivariate functional principal component analysis relies on a univariate basis expansion for each element $X^{(j)}$.
    - The univariate basis representation is calculated using the univDecomp function, that passes the univariate functional observations and optional parameters to the specific function.
    - The univariate decompositions are specified via the uniExpansions argument in the MFPCA function.
    - It is a list of the same length as the mFData object, i.e. having one entry for each element of the multivariate functional data.
    - For each element, uniExpansion must specify at least the type of basis functions to use.
    - Additionally, one may add further parameters.
    - The following basis representations are supported: \itemize{
        -  Given basis functions. Then \code{uniExpansions[[j]] = list(type = "given", functions)}, where \code{functions} is a \code{funData} object on the same domain as \code{mFData}, containing the given basis functions. 
        - Univariate functional principal component analysis. Then \code{uniExpansions[[j]] = list(type = "uFPCA", nbasis, pve, npc, makePD)}, where \code{nbasis,pve,npc,makePD} are parameters passed to the \code{\link{PACE}} function for calculating the univariate functional principal component analysis.
        - Basis functions expansions from the
  package \pkg{fda}. Then \code{uniExpansions[[j]] = list(type = "fda", ...)},
#' where \code{...} are passed to \code{\link[funData]{funData2fd}}, which
#' heavily builds on \code{\link[fda]{eval.fd}}. If \pkg{fda} is not available,
#' a warning is thrown.
\item Spline basis functions (not penalized). Then
#' \code{uniExpansions[[j]] = list(type = "splines1D", bs, m, k)}, where
#' \code{bs,m,k} are passed to the functions \code{\link{univDecomp}} and
#' \code{\link{univExpansion}}. For two-dimensional tensor product splines, use
#' \code{type = "splines2D"}. 
\item Spline basis functions (with smoothness
#' penalty). Then \code{uniExpansions[[j]] = list(type = "splines1Dpen", bs, m,
#' k)}, where \code{bs,m,k} are passed to the functions \code{\link{univDecomp}}
#' and \code{\link{univExpansion}}. Analogously to the unpenalized case, use
#' \code{type = "splines2Dpen"} for 2D penalized tensor product splines. 
\item
#' Cosine basis functions. Use \code{uniExpansions[[j]] = list(type = "DCT2D",
#' qThresh, parallel)} for functions one two-dimensional domains (images) and
#' \code{type = "DCT3D"} for 3D images. The calculation is based on the discrete
#' cosine transform (DCT) implemented in the C-library \code{fftw3}. If this
#' library is not available, the function will throw  a warning. \code{qThresh}
#' gives the quantile for hard thresholding the basis coefficients based on
#' their absolute value. If \code{parallel = TRUE}, the coefficients for
#' different images are calculated in parallel.} See \code{\link{univDecomp}}
#' and \code{\link{univExpansion}} for details.}

@@

########### modified by Jongmin Mun #############################
  # calculate univariate basis expansion for all components
  uniBasis <- mapply(
    function(expansion, data, data_predict){
      do.call(univDecomp, c(
        list(funDataObject = data, funDataObject_predict = data_predict),
        expansion)
        )
      },
    expansion = uniExpansions,
    data = mFData,
    data_predict = mFData_predict,
    SIMPLIFY = FALSE
    )
########### modified by Jongmin Mun #############################

@@

@@

@@

```{r}
library(refund)
eval_point<-seq(1, 100, length.out = 100)
n_iter_max = 40
n_basis_predictor = 10
n_scalar <- 9
library(foreach)
library(funData)
uniExpansions <- list(
  list(type = "uFPCA", nbasis = n_basis_predictor, npc = n_iter_max),
  list(type = "uFPCA", nbasis = n_basis_predictor, npc = n_iter_max)
)

>>>>

The class definition is a simple list that contains a matrix and a collection of \code{fd} objects from the \code{fda} package, along with the necessary metadata.
- gram matrices are saved as a list.
```{r classdef:predictor_hybrid}
#' Create a predictor_hybrid object (S3 version, automatic basis size and Gram matrix)
#' Constructs a hybrid predictor object that stores both scalar and functional predictors,
#' and automatically computes Gram matrices (inner products of basis functions).
#'
#' @param Z A numeric matrix of dimension \code{n_sample × n_scalar} representing scalar predictors.
#' @param functional_list A list of functional predictors (e.g., \code{fd} objects from the \code{fda} package).
#'
#' @return An object of class \code{predictor_hybrid}, containing scalar and functional data with Gram matrices.
#' @export
predictor_hybrid <- function(Z, functional_list, eval_point) {
  stopifnot(is.matrix(Z), is.numeric(Z))
  stopifnot(is.list(functional_list))

@@

  structure(
    list(
      Z = Z,
      functional_list = functional_list,
      eval_point = eval_point,
      gram_list = gram_list,
      gram_deriv2_list = gram_deriv2_list, # Store the new list
      n_basis_list = n_basis_list,
      n_sample = n_sample,
      n_functional = n_functional,
      n_scalar = n_scalar
    ),
    class = "predictor_hybrid"
  )
}
```

@@

@@

## Modified MFPCA Function for Simultaneous Training and Prediction

The MFPCA (Multivariate Functional Principal Component Analysis) function computes multivariate functional PCA for functions defined across different (dimensional) domains.
We have modified the original implementation to admit a new input, **`mFData_predict`**, allowing the function to calculate PCA scores for **both the training data and new test (or validation) data simultaneously** in a single run. This change was necessary because the standard `predict.MFPCAfit` function does not permit computing scores for new, out-of-sample functions.
The multivariate functional principal component analysis inherently relies on a **univariate basis expansion** for each functional element ($X^{(j)}$). To simplify the code modification, we use only the univariate functional PCA option and have removed other options provided by the original MFPCA function. We modifies the main function and two internal functions.

### MFPCA

@@

########### Below is modified by Jongmin Mun #############################
  # calculate univariate basis expansion for all components
  uniBasis <- mapply(
    function(expansion, data, data_predict){
      do.call(univDecomp, c(
        list(funDataObject = data, funDataObject_predict = data_predict),
        expansion)
        )
      },
    expansion = uniExpansions,
    data = mFData,
    data_predict = mFData_predict,
    SIMPLIFY = FALSE
    )
########### Above is modified by Jongmin Mun #############################

@@

### fpcaBasis

```{r}
################# modified by Jongmin Mun #####################
fpcaBasis <- function(funDataObject, funDataObject_predict, nbasis = 10, pve = 0.99, npc = NULL, makePD = FALSE, cov.weight.type = "none")
{
  FPCA <- PACE(funDataObject, predData = NULL, nbasis, pve, npc, makePD, cov.weight.type)
  FPCA_pred <- PACE(funDataObject = funDataObject, predData = funDataObject_predict, nbasis, pve, npc, makePD, cov.weight.type)
  return(list(scores = FPCA$scores,
              scores.pred = FPCA_pred$scores,
              ortho = TRUE,
              functions = FPCA$functions,
              meanFunction = FPCA$mu
  ))
}
```

### univDecomp
```{r}
univDecomp <- function(type, funDataObject, funDataObject_predict, ...)
{
  # Parameter checking
  if(is.null(type))
    stop("Parameter 'type' is missing.")
  else
  {
    if(!is.character(type))
      stop("Parameter 'type' must be a character string. See ?univDecomp for details.")
  }

  if(class(funDataObject) != "funData")
    stop("Parameter 'funDataObject' must be a funData object.")

  # get all arguments (except for function call and type)
  params <- list(...)

  # check if type and data are of correct type
  if(is.null(type))
    stop("univDecomp: must specify 'type'.")

  if(!inherits(type, "character"))
    stop("univDecomp: 'type' must be of class character.")

  if(is.null(funDataObject))
    stop("univDecomp: must specify 'funDataObject'.")

  if(class(funDataObject) != "funData")
    stop("univDecomp: 'funDataObject' must be of class funData.")

  params$funDataObject <- funDataObject # add funDataObject (-> make sure is evaluated in correct env.)

  ############### modified by Jongmin Mun ######################
  if (type == "uFPCA"){params$funDataObject_predict <- funDataObject_predict}
  ############### modified by Jongmin Mun ######################

  res <- switch(type,
                "given" = do.call(givenBasis, params),
                "uFPCA" = do.call(fpcaBasis, params),
                "UMPCA" = do.call(umpcaBasis, params),
                "FCP_TPA" = do.call(fcptpaBasis, params),
                "splines1D" = do.call(splineBasis1D, params),
                "splines1Dpen" = do.call(splineBasis1Dpen, params),
                "splines2D" = do.call(splineBasis2D, params),
                "splines2Dpen" = do.call(splineBasis2Dpen, params),
                "fda" = do.call(fdaBasis, params),
                "DCT2D" = do.call(dctBasis2D, params),
                "DCT3D" = do.call(dctBasis3D, params),
                stop("Univariate Decomposition for 'type' = ", type, " not defined!")
  )

  if(res$ortho == FALSE & is.null(res$B))
    stop("UnivDecomp: must provide integral matrix B for non-orthonormal basis functions.")

## Unmodified functions

### calcBasisIntegrals
```{r}

@@

### .PACE

```{r}
.PACE <- function(X, Y, Y.pred = NULL, nbasis = 10, pve = 0.99, npc = NULL, makePD = FALSE, cov.weight.type = "none")
{
  if (is.null(Y.pred))
    Y.pred = Y
  D = NCOL(Y)
  if(D != length(X)) # check if number of observation points in X & Y are identical
    stop("different number of (potential) observation points differs in X and Y!")
  I = NROW(Y)
  I.pred = NROW(Y.pred)
  d.vec = rep(X, each = I) # use given X-values for estimation of mu
  gam0 = mgcv::gam(as.vector(Y) ~ s(d.vec, k = nbasis))
  mu = mgcv::predict.gam(gam0, newdata = data.frame(d.vec = X))
  Y.tilde = Y - matrix(mu, I, D, byrow = TRUE)
  cov.sum = cov.count = cov.mean = matrix(0, D, D)
  for (i in seq_len(I)) {
    obs.points = which(!is.na(Y[i, ]))
    cov.count[obs.points, obs.points] = cov.count[obs.points, obs.points] + 1
    cov.sum[obs.points, obs.points] = cov.sum[obs.points, obs.points] + tcrossprod(Y.tilde[i, obs.points])
  }
  G.0 = ifelse(cov.count == 0, NA, cov.sum/cov.count)
  diag.G0 = diag(G.0)
  diag(G.0) = NA
  row.vec = rep(X, each = D) # use given X-values
  col.vec = rep(X, D) # use given X-values
  cov.weights <- switch(cov.weight.type,
                        none = rep(1, D^2),
                        counts = as.vector(cov.count),
                        stop("cov.weight.type ", cov.weight.type, " unknown in smooth covariance estimation"))

  npc.0 = matrix(mgcv::predict.gam(mgcv::gam(as.vector(G.0)~te(row.vec, col.vec, k = nbasis), weights = cov.weights),
                                   newdata = data.frame(row.vec = row.vec, col.vec = col.vec)), D, D)
  npc.0 = (npc.0 + t(npc.0))/2
  # no extra-option (useSymm) as in fpca.sc-method
  if (makePD) { # see fpca.sc
    npc.0 <- {
      tmp <- Matrix::nearPD(npc.0, corr = FALSE, keepDiag = FALSE,
                            do2eigen = TRUE, trace = options()$verbose)
      as.matrix(tmp$mat)
    }
  }

  ### numerical integration for calculation of eigenvalues (see Ramsay & Silverman, Chapter 8)
  w <- funData::.intWeights(X, method = "trapezoidal")
  Wsqrt <- diag(sqrt(w))
  Winvsqrt <- diag(1/(sqrt(w)))
  V <- Wsqrt %*% npc.0 %*% Wsqrt
  evalues = eigen(V, symmetric = TRUE, only.values = TRUE)$values
  ###
  evalues = replace(evalues, which(evalues <= 0), 0)
  npc = ifelse(is.null(npc), min(which(cumsum(evalues)/sum(evalues) > pve)), npc)
  efunctions = matrix(Winvsqrt%*%eigen(V, symmetric = TRUE)$vectors[, seq(len = npc)], nrow = D, ncol = npc)
  evalues = eigen(V, symmetric = TRUE, only.values = TRUE)$values[seq_len(npc)]  # use correct matrix for eigenvalue problem
  cov.hat = efunctions %*% tcrossprod(diag(evalues, nrow = npc, ncol = npc), efunctions)
  ### numerical integration for estimation of sigma2
  T.len <- X[D] - X[1] # total interval length
  T1.min <- min(which(X >= X[1] + 0.25*T.len)) # left bound of narrower interval T1
  T1.max <- max(which(X <= X[D] - 0.25*T.len)) # right bound of narrower interval T1
  DIAG = (diag.G0 - diag(cov.hat))[T1.min :T1.max] # function values
  # weights
  w <- funData::.intWeights(X[T1.min:T1.max], method = "trapezoidal")
  sigma2 <- max(1/(X[T1.max]-X[T1.min]) * sum(DIAG*w, na.rm = TRUE), 0) #max(1/T.len * sum(DIAG*w), 0)
  ####
  D.inv = diag(1/evalues, nrow = npc, ncol = npc)
  Z = efunctions
  Y.tilde = Y.pred - matrix(mu, I.pred, D, byrow = TRUE)
  fit = matrix(0, nrow = I.pred, ncol = D)
  scores = matrix(NA, nrow = I.pred, ncol = npc)
  # no calculation of confidence bands, no variance matrix
  for (i.subj in seq_len(I.pred)) {
    obs.points = which(!is.na(Y.pred[i.subj, ]))
    if (sigma2 == 0 & length(obs.points) < npc) {
      stop("Measurement error estimated to be zero and there are fewer observed points than PCs; scores cannot be estimated.")
    }
    Zcur = matrix(Z[obs.points, ], nrow = length(obs.points),
                  ncol = dim(Z)[2])
    ZtZ_sD.inv = solve(crossprod(Zcur) + sigma2 * D.inv)
    scores[i.subj, ] = ZtZ_sD.inv %*% crossprod(Zcur, Y.tilde[i.subj, obs.points])
    fit[i.subj, ] = t(as.matrix(mu)) + tcrossprod(scores[i.subj, ], efunctions)
  }
  ret.objects = c("fit", "scores", "mu", "efunctions", "evalues",
                  "npc", "sigma2") # add sigma2 to output
  ret = lapply(seq_len(length(ret.objects)), function(u) get(ret.objects[u]))
  names(ret) = ret.objects
  ret$estVar <- diag(cov.hat)
  return(ret)
}
```

### PACE

```{r}
PACE <- function(funDataObject, predData = NULL, nbasis = 10, pve = 0.99, npc = NULL, makePD = FALSE, cov.weight.type = "none")
{
  # check inputs
  if(! class(funDataObject) %in% c("funData", "irregFunData"))
    stop("Parameter 'funDataObject' must be a funData or irregFunData object.")
  if(dimSupp(funDataObject) != 1)
    stop("PACE: Implemented only for funData objects with one-dimensional support.")
  if(methods::is(funDataObject, "irregFunData")) # for irregular functional data, use funData representation
    funDataObject <- as.funData(funDataObject)

  if(is.null(predData))
    Y.pred = NULL # use only funDataObject
  else
  {
    if(!isTRUE(all.equal(funDataObject@argvals, predData@argvals)))
      stop("PACE: funDataObject and predData must be defined on the same domains!")

    Y.pred = predData@X
  }

  if(!all(is.numeric(nbasis), length(nbasis) == 1, nbasis > 0))
    stop("Parameter 'nbasis' must be passed as a number > 0.")

  if(!all(is.numeric(pve), length(pve) == 1, 0 <= pve, pve <= 1))
    stop("Parameter 'pve' must be passed as a number between 0 and 1.")

  if(!is.null(npc) & !all(is.numeric(npc), length(npc) == 1, npc > 0))
    stop("Parameter 'npc' must be either NULL or passed as a number > 0.")

  if(!is.logical(makePD))
    stop("Parameter 'makePD' must be passed as a logical.")

  if(!is.character(cov.weight.type))
    stop("Parameter 'cov.weight.type' must be passed as a character.")

  res <- .PACE(X = funDataObject@argvals[[1]],
               Y = funDataObject@X,
               Y.pred = Y.pred,
               nbasis = nbasis, pve = pve, npc = npc, makePD = makePD,
               cov.weight.type = cov.weight.type)

  return(list(mu = funData(funDataObject@argvals, matrix(res$mu, nrow = 1)),
              values = res$evalues,
              functions = funData(funDataObject@argvals, t(res$efunctions)),
              scores = res$scores,
              scores.pred = res$scores.pred,
              fit = funData(funDataObject@argvals, res$fit),
              npc = res$npc,
              sigma2 = res$sigma2,
              estVar = funData(funDataObject@argvals, matrix(res$estVar, nrow = 1))
  ))
}
```

### expandBasisFunction

```{r}
expandBasisFunction <- function(scores, argvals = functions@argvals, functions)
{
  if(dim(scores)[2] != nObs(functions))
    stop("expandBasisFunction: number of scores for each observation and number of eigenfunctions does not match.")

  # collapse higher-dimensional functions, multiply with scores and resize the result
  d <- dim(functions@X)
  nd <- length(d)

  if(nd == 2)
    resX <- scores %*% functions@X

  if(nd == 3)
  {
    resX <- array(NA, dim = c(dim(scores)[1], d[-1]))

    for(i in seq_len(d[2]))
      resX[,i,] <- scores %*% functions@X[,i,]
  }

  if(nd == 4)
  {
    resX <- array(NA, dim = c(dim(scores)[1], d[-1]))

    for(i in seq_len(d[2]))
      for(j in seq_len(d[3]))
          resX[,i,j,] <- scores %*% functions@X[,i,j,]
  }

  if(nd > 4) # slow solution due to aperm
  {
    resX <- aperm(plyr::aaply(.data = functions@X, .margins = 3:nd, 
                        .fun = function(x,y){y %*% x}, y = scores), 
                  c(nd-1,nd, seq_len((nd-2))))
    dimnames(resX) <- NULL
  }

  return( funData(argvals, resX) )
}
```

### univExpansion

```{r}
univExpansion <- function(type, scores, argvals = ifelse(!is.null(functions), functions@argvals, NULL), functions, params = NULL)
{
  # Parameter checking
  if(is.null(type))
    stop("Parameter 'type' is missing.")
  else
 {
   if(!is.character(type))
     stop("Parameter 'type' must be a character string. See ?univExpansion for details.")
 }

  if(is.null(scores))
    stop("Parameter 'scores' is missing.")
  else
    {
      if(!is.matrix(scores))
      stop("Parameter 'scores' must be passed as a matrix.")
    }

  if(is.numeric(argvals))
  {
    argvals <- list(argvals)
    warning("Parameter 'argvals' was passed as a vector and transformed to a list.")
  }

  if(is.null(functions))
  {
    if(is.null(argvals))
      stop("Must pass 'argvals' if 'functions' is NULL.")
    else
      {
        if(!is.list(argvals)) 
          stop("Parameter 'argvals' must be passed as a list.")
      }
  }
  else
  {
    if(class(functions) != "funData")
      stop("Parameter 'functions' must be a funData object.")

    # check interaction with other parameters
    if(nObs(functions) != NCOL(scores))
      stop("Number of scores per curve does not match the number of basis functions.")

    if(!is.null(argvals) & !isTRUE(all.equal(argvals, functions@argvals)))
      stop("The parameter 'argvals' does not match the argument values of 'functions'.")
  }

  if(!is.null(params) & !is.list(params))
    stop("The parameter 'params' must be passed as a list.")

  # start calculations
  params$scores <- scores
  params$functions <- functions

  if(is.numeric(argvals))
    argvals <- list(argvals)

  params$argvals <- argvals

  res <- switch(type,
                "given" = do.call(expandBasisFunction, params),
                "uFPCA" = do.call(expandBasisFunction, params),
                "UMPCA" = do.call(expandBasisFunction, params),
                "FCP_TPA" = do.call(expandBasisFunction, params),
                "splines1D" = do.call(splineFunction1D, params),
                "splines1Dpen" = do.call(splineFunction1D, params),
                "splines2D" = do.call(splineFunction2D, params),
                "splines2Dpen" = do.call(splineFunction2Dpen, params),
                "fda" = do.call(expandBasisFunction, params),
                "DCT2D" = do.call(dctFunction2D, params),
                "DCT3D" = do.call(dctFunction3D, params),
                "default" = do.call(expandBasisFunction, params),
                stop("Univariate Expansion for 'type' = ", type, " not defined!")
  )

## end-to-end MFPCA regression method

```{r}
# --- 1. MAIN FUNCTION: fit.mfpca_regression ---

#' End-to-End MFPCA Regression Pipeline
#'
#' Performs sample splitting, normalization, cross-validation (CV) to select
#' optimal functional (FPC) and scalar (SPC) components, linear regression fitting,
#' and test set prediction.
#'
#' @param W predictor_hybrid object containing all predictors.
#' @param y Vector. Response variable.
#' @param train_ratio Numeric scalar between 0 and 1 for splitting data (default 0.7).
#' @param n_fold Integer. Number of folds for cross-validation (default 5).
#' @param n_fpc_max Integer. Max number of FPCs to consider in CV (default 10).
#' @param n_spc_max Integer. Max number of SPCs to consider in CV (default 5).
#' @return List containing the final model fit, optimal parameters, and test MSE.
#' @export
fit.mfpca_regression <- function(W, y, train_ratio = 0.7, n_fold = 5, n_fpc_max = 10, n_spc_max = 5) {

  # --- 0. Setup and Initial Checks ---
  stopifnot(inherits(W, "predictor_hybrid"))

  # Standard parameters
  eval_point <- seq(0, 1, length.out = 100)

  uniExpansions <- list(
    list(type = "uFPCA", nbasis = 10, npc = n_fpc_max), 
    list(type = "uFPCA", nbasis = 10, npc = n_fpc_max)
  )

  # Adjust max components
  n_scalar <- W$n_scalar
  if (n_spc_max > n_scalar) n_spc_max <- n_scalar

  # --- 1. Split and Normalize Full Data ---

  split_obj <- split.hybrid(W, y, train_ratio)
  split_obj_normalized <- normailize.all(split_obj)

  # Extract components
  y_train_norm <- split_obj_normalized$response_train
  Z_train_norm <- split_obj_normalized$data_scalar_train
  f1_train_norm <- split_obj_normalized$data_fun_1_train
  f2_train_norm <- split_obj_normalized$data_fun_2_train
  N_train <- length(y_train_norm)

  # --- 2. Cross-Validation Setup ---

  cv_mat <- matrix(Inf, nrow = n_fpc_max, ncol = n_spc_max) 
  kfold_idx <- create_idx_kfold(N_train, n_fold)

  # --- 3. Cross-Validation Loop (Find Best n_fpc, n_spc) ---

  for (fold_num in 1:n_fold) {
    valid_idx <- kfold_idx[[fold_num]]
    train_idx_cv <- (1:N_train)[-valid_idx]

    # 3a. Prepare CV training and validation sets
    f1_train_cv <- f1_train_norm[train_idx_cv]
    f2_train_cv <- f2_train_norm[train_idx_cv]
    f1_valid_cv <- f1_train_norm[valid_idx]
    f2_valid_cv <- f2_train_norm[valid_idx]

    # Format functional data
    predictor_cv_train <- multiFunData(
      fd2funData(f1_train_cv, eval_point),
      fd2funData(f2_train_cv, eval_point)
    )
    predictor_cv_valid <- multiFunData(
      fd2funData(f1_valid_cv, eval_point),
      fd2funData(f2_valid_cv, eval_point)
    )

    Z_train_cv <- Z_train_norm[train_idx_cv, , drop = FALSE]
    Z_valid_cv <- Z_train_norm[valid_idx, , drop = FALSE]
    y_train_cv <- y_train_norm[train_idx_cv]
    y_valid_cv <- y_train_norm[valid_idx]

    # 3b. Fit PCA Models on CV Training Data
    MFPCA_fit_cv <- MFPCA(mFData = predictor_cv_train, mFData_predict = predictor_cv_valid, 
                          M = n_fpc_max, uniExpansions = uniExpansions)
    PCA_fit_cv <- prcomp(Z_train_cv, center = FALSE, scale = FALSE)

    FPC_scores_train_cv <- MFPCA_fit_cv$scores
    FPC_scores_valid_cv <- MFPCA_fit_cv$scores.pred
    PC_scores_train_cv <- PCA_fit_cv$x
    PC_scores_valid_cv <- predict(PCA_fit_cv, Z_valid_cv)

    # 3c. Inner Parameter Loop (Fit and Validate)
    for (n_fpc in 1:n_fpc_max) {
      for (n_spc in 1:n_spc_max) {

        # Select components
        FPC_train_select <- FPC_scores_train_cv[, 1:n_fpc, drop = FALSE]
        FPC_valid_select <- FPC_scores_valid_cv[, 1:n_fpc, drop = FALSE]
        PC_train_select <- PC_scores_train_cv[, 1:n_spc, drop = FALSE]
        PC_valid_select <- PC_scores_valid_cv[, 1:n_spc, drop = FALSE]

        # Combine selected scores for linear model
        train_data_lm <- data.frame(response = y_train_cv, FPC_train_select, PC_train_select)
        valid_data_lm <- data.frame(FPC_valid_select, PC_valid_select)

        # Fit Linear Regression Model
        linear_fit_cv <- lm(response ~ ., data = train_data_lm)

        # Predict, Calculate MSE, and Accumulate
        y_pred_valid <- predict(linear_fit_cv, valid_data_lm)
        cv_mat[n_fpc, n_spc] <- cv_mat[n_fpc, n_spc] + mean((y_pred_valid - y_valid_cv)^2)
      }
    }
  }

  # 4. Select Optimal Parameters
  min_idx <- arrayInd(which.min(cv_mat), dim(cv_mat))
  n_fpc_best <- min_idx[1]
  n_spc_best <- min_idx[2]

  # --- 5. Final Model Training and Prediction (Full Training Set) ---

  # Extract final components
  Z_test_norm <- split_obj_normalized$data_scalar_test
  f1_test_norm <- split_obj_normalized$data_fun_1_test
  f2_test_norm <- split_obj_normalized$data_fun_2_test
  y_test_norm <- split_obj_normalized$response_test

  # Prepare final functional data
  predictor_final_train <- multiFunData(fd2funData(f1_train_norm, eval_point), fd2funData(f2_train_norm, eval_point))
  predictor_final_test <- multiFunData(fd2funData(f1_test_norm, eval_point), fd2funData(f2_test_norm, eval_point))

  # Fit final PCA models
  MFPCA_fit_final <- MFPCA(mFData = predictor_final_train, mFData_predict = predictor_final_test, 
                          M = n_fpc_best, uniExpansions = uniExpansions)
  PCA_fit_final <- prcomp(Z_train_norm, center = FALSE, scale = FALSE)

  # Extract and combine final scores
  FPC_train_final <- MFPCA_fit_final$scores
  FPC_test_final <- MFPCA_fit_final$scores.pred
  PC_train_final <- PCA_fit_final$x[, 1:n_spc_best, drop = FALSE]
  PC_test_final <- predict(PCA_fit_final, Z_test_norm)[, 1:n_spc_best, drop = FALSE]

  train_data_final_lm <- data.frame(response = y_train_norm, FPC_train_final, PC_train_final)
  test_data_final_lm <- data.frame(FPC_test_final, PC_test_final)

  # Fit Final Linear Model
  final_linear_fit <- lm(response ~ ., data = train_data_final_lm)

  # Predict and Evaluate
  y_pred_test <- predict(final_linear_fit, test_data_final_lm)
  final_test_rmse <- sqrt(mean((y_pred_test - y_test_norm)^2))

  # --- 6. Return Results ---
  return(list(
    n_fpc_best = n_fpc_best,
    n_spc_best = n_spc_best,
    test_rmse = final_test_rmse,
    final_model = final_linear_fit,
    regression_coefficients = coef(final_linear_fit)
  ))
}
```

@@

```{r}
library(refund)
eval_point<-seq(1, 100, length.out = 100)/100
n_iter_max = 40
n_basis_predictor = 10
n_scalar <- 9
library(foreach)
library(funData)
f1 = fd2funData(create_mock_fd(N, 1), eval_point)
f2 = fd2funData(create_mock_fd(N, 2), eval_point)
uniExpansions <- list(
  list(type = "uFPCA", nbasis = n_basis_predictor, npc = n_iter_max),
  list(type = "uFPCA", nbasis = n_basis_predictor, npc = n_iter_max)
)
